---
title: "Banding processing"
author: "LGCarlson"
date: "7/29/2021"
output: html_document
---

## Banding codes

_age:_

* 0 unknown

* 1 AHY

* 2 HY

* 3 Juvenile

* 4 Local

* 5 SY

* 6 ASY

_sex:_

* 0 unknown

* 4 M

* 5 F

_b_dir:_

* 0 NW quarter of the earth

_b_flyway:_

* 2 Mississippi

* 3 Central

* 6 Canada

_b_region:_

[codes here](https://www.pwrc.usgs.gov/BBL/MANUAL/reg.cfm#region)
Canadian categories to keep: (US Flyways okay)

* 43 Northwest Territories

* 40 Nunavut

* 04 Alberta

* 79 Saskatchewan

* 45 Manitoba

* 68 Ontario

_b_day_code:_

* 0 unknown

* 1 Day 1-5

* 2 Day 6-10

* 3 Day 11-15

* 4 Day 16-20

* 5 Day 21-25

* 6 Day >26

_status:_

* 2 transported

* 3 normal wild

* 4 hand reared

* 5 sick, injured, deformity

* 6,7,8 experimental, etc

_sp_num:_

* 1320 MALL

_ai_vai:_

Categories to keep:

* Federal numbered metal band only. 

* Captured by spotlighting. 

* Misc. metal band (State, Provincial etc) with address or telephone number, plus Federal band.

* Miscellaneous band, Federal band, plus auxiliary marker(s). 

## Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(lubridate)
library(janitor)
library(rnaturalearth)
library(rnaturalearthdata)
library(ggspatial)
library(sf)
```

```{r}
theme_set(theme_light())

#set personalized plotting theme
LGCtheme <- theme_light() + theme(panel.background = element_blank(), 
                panel.grid = element_blank(), 
                axis.text = element_text(size = 13), 
                axis.title = element_text(size = 14), 
                legend.text = element_text(size = 13), 
                legend.title = element_text(size = 14), 
                strip.background = element_blank(), 
                strip.text = element_text(size = 13, color = "black"))

gmri_qual_extended<-c("#003A54","#00608A","#5D96AF","#3F4B24","#6F7E4B","#97AA5C","#A8B212","#AF9700","#EACA00","#FFE961","#BB3F0E", "#EA4F12","#F4916B","#00736D", "#55A29E","#A0CDCB","#767676","#A7A7A7","#E2E2E2")
```


```{r}
# Table to convert date range to center 
date_code_tib<-tribble(
  ~b_day_code, ~b_day, 
  1, 5, 
  2, 10, 
  3, 15, 
  4, 20, 
  5, 25, 
  6, 28)

# Midcontinent Canadian provinces
provs_keep<-c(43,40,04,79,45,68)

# Band/capture types to keep
ai_vai_vec<-c("Federal numbered metal band only.","Captured by spotlighting.", "Misc. metal band (State, Provincial etc) with address or telephone number, plus Federal band.","Miscellaneous band, Federal band, plus auxiliary marker(s).")

# Region (state/province) codes and names
b_region_codes<-read_csv(here::here("raw_data", "region_codes.csv")) %>% 
  filter(!is.na(region_abbrev)) %>% 
  dplyr::select(-region_name,-flyway_code) %>% 
  dplyr::rename("b_region" = "region_code","b_region_abbrev" = "region_abbrev")

```

## Load data

```{r, echo=FALSE, message=F}
# Load in Canada banding records
ca_band<-read_csv(here::here("raw_data", "canada_banding.csv")) %>% 
  janitor::clean_names() %>% 
  mutate(b_flyway_name = "Canada") %>% 
  filter(b_region %in% provs_keep)
      #keep only data from birds banded in mid-continent Canadian provinces

# Load in Central banding records
cl_band<-read_csv(here::here("raw_data", "central_banding.csv")) %>% 
  janitor::clean_names() %>% 
  mutate(b_flyway_name = "Central")

# Load in Mississippi banding records
ms_band<-read_csv(here::here("raw_data", "miss_banding.csv")) %>% 
  janitor::clean_names() %>% 
  mutate(b_flyway_name = "Mississippi")
```


_Modifications:_

* join all banding records 

* remove indv. where banding month is unknown

* remove indv. where banding day is unknown

* join date_code_tib by b_day_code, so that b_day so that approximate banding_dates can be created using maximum dates in b_day_code range

* create a new column with banding date


```{r}
all_bands<-bind_rows(ca_band, cl_band, ms_band) %>% 
  filter(b_month <= 12, 
         b_day_code > 0) %>% 
  left_join(date_code_tib, by = "b_day_code") %>% 
  mutate(b_date = as.Date(paste(b_year,b_month,b_day, sep = "-"))) %>% 
  left_join(b_region_codes, by = "b_region")
```

## Check data

```{r}
table(all_bands$age)

# will remove all but 4 and 5
table(all_bands$sex)

# will need to remove (if unknown) or specify (approx known) non-real months
table(all_bands$b_month)


# will need to remove (if unknown) or specify (approx known) non-real days 
table(all_bands$b_day_code)

table(all_bands$status)

table(all_bands$b_region_abbrev)

#table(all_bands$ai_vai)
```

## Processing code

_Modifications(in order of operation):_

* limit years to 1950-2020

* keep only pre-season banded birds (banding months 6-9)

* keep only normal, wild birds (status = 3)

* keep only known sexes (sex = 4,5)

* keep only desired ai_vai's

* create unique id 

* re-order columns manually

```{r}
final_bands<-all_bands %>% 
  filter(b_year >=1950 & b_year <=2020) %>% 
  filter(b_month >=6 & b_month <=9) %>% 
  filter(status == 3) %>% 
  filter(sex == 4 | sex == 5) %>% 
  filter(ai_vai %in% ai_vai_vec) %>% 
  mutate(id = paste(band_prefix_plus,count_of_birds,age,sex,b_date,gisb_lat,gisb_long,ai_vai,how_sexed,sep="_"))%>%
  dplyr::select(band_prefix_plus,count_of_birds,age,age_vage,sex,sp_num,status,b_flyway,b_flyway_name,b_date,b_year,b_month,b_day_code, b_day, country_code,b_region,b_region_abbrev,state_code,b_dir,b_coord_prec,b_lat,b_long,
                gisb_lat,gisb_long, band_size,band_type,ai_vai,how_aged,how_sexed,id) %>% 
  distinct(id, .keep_all = T)


#confirm that "id" field works as a unique identifier
length(final_bands$id)
length(unique(final_bands$id))
```

## Spatial processing

```{r, message=F, warning=F, echo=F}
# load in world and state sf
world <- rnaturalearth::ne_countries(scale = "medium", returnclass = "sf")

states <- rnaturalearth::ne_states(country = c("canada", "united states of america"), returnclass = "sf")


#load in flyway shapefiles
cent_flyway <- sf::st_read(here::here("shapefiles","Central_Flyway.shp")) %>% 
  st_transform(crs = "+init=epsg:4326") 

miss_flyway <- sf::st_read(here::here("shapefiles","Mississippi_Flyway.shp")) %>% 
  st_transform(crs = 4326) 

pac_flyway <- sf::st_read(here::here("shapefiles","Pacific_Flyway.shp")) %>% 
  st_transform(crs = 4326) 
```

```{r}
# Create a summary dataframe for banding locations
banding_locs<-final_bands %>% 
  group_by(gisb_lat,gisb_long) %>% 
  summarise(n_bands=n())
```

Plot banding locations. Will need to remove birds banded in the Pacific flyway (outside mid-continent). 

```{r, fig.height=8, fig.width=8}
ggplot(data = world) + geom_sf(fill="white") + 
  geom_sf(data=states, fill=NA) + 
  geom_sf(data=cent_flyway, fill="#6F7E4B") +
  geom_sf(data=miss_flyway, fill="#00608A") + 
  geom_sf(data=pac_flyway, fill="#F4916B") +
  geom_point(data=banding_locs, aes(x=gisb_long, y=gisb_lat, size=n_bands)) +
  coord_sf(xlim = c(-130, -70), ylim = c(25, 70), expand = FALSE) + 
  labs(x="Banding longitude", y="Banding latitude", size="Number banded")
```


```{r}
# Create spatial data
final_bands_sf<-st_as_sf(final_bands, coords = c("gisb_long","gisb_lat"), crs = 4326) 
class(final_bands_sf)

length(unique(final_bands_sf$id))
```

Remove encounters of birds that were *banded* in Pacific flyway.
```{r}
# This selects all individuals that were banded inside the Pacific flyway shapefile
points_inpac <- final_bands_sf[pac_flyway, ]

# Creates a vector of Pacific flyway banded individuals
to_remove<-c(points_inpac$id)


# Creates dataframe of Pacific flyway banded individuals to plot and check for correctness
final_bands_inpac<-final_bands %>% 
  filter(id %in% to_remove) %>% 
   

# Plot to confirm
ggplot(data = world) + geom_sf(fill="white") + 
  geom_sf(data=states, fill=NA) + 
  geom_sf(data=pac_flyway, fill="#F4916B") +
  geom_point(data=final_bands_inpac, aes(x=gisb_long, y=gisb_lat)) +
  coord_sf(xlim = c(-130, -70), ylim = c(25, 70), expand = FALSE) + 
  labs(x="Banding longitude", y="Banding latitude")
```


```{r}
# Create custom "not in" function
`%not_in%` <- purrr::negate(`%in%`)

# Remove individuals banded in Pacific flyway
midcontinent_bands<-final_bands %>% 
  filter(id %not_in% to_remove)

# Plot to confirm
ggplot(data = world) + geom_sf(fill="white") + 
  geom_sf(data=states, fill=NA) + 
  geom_sf(data=cent_flyway, fill="#6F7E4B") +
  geom_sf(data=miss_flyway, fill="#00608A") + 
  geom_point(data=midcontinent_bands, aes(x=gisb_long, y=gisb_lat)) +
  coord_sf(xlim = c(-130, -70), ylim = c(25, 70), expand = FALSE) + 
  labs(x="Banding longitude", y="Banding latitude")

```


## Save new datafile 

```{r}
## write_csv(midcontinent_bands, here::here("proc_data","banding_data.csv"))
```

```{r}
midcontinent_bands %>% 
  summarise(n=sum(count_of_birds))
```


## Visualize data

Majority of encountered individuals banded in the midcontinent were banded in Canada.
More males banded than females in Canada, fairly even in USA. 

```{r, fig.height=4, fig.width=10}
midcontinent_bands %>% 
  group_by(b_year, sex, b_flyway_name) %>% 
  summarise(total_banded = sum(count_of_birds)) %>% 
  ggplot(aes(x=b_year, y=total_banded, color=as.factor(sex))) + geom_line() + 
  facet_wrap(~b_flyway_name, ncol=3) + 
  scale_color_manual(values=c("#5D96AF","#C07F9C")) + LGCtheme + theme(legend.position = "none") +
  labs(x="Banding Year", y="Total banded", title="MALLs banded in each flyway, by banding year, by sex")
```



Banding was restricted to pre-season, but most preseason banding occurs in August and September.

```{r, fig.height=4.5, fig.width=7.5}
midcontinent_bands %>% 
  mutate(b_month = as.factor(b_month)) %>% 
  group_by(b_year, b_month) %>% 
  summarise(total_banded = sum(count_of_birds)) %>% 
  ggplot(aes(x=b_year, y=total_banded, color=forcats::fct_reorder(b_month, total_banded, .desc = T))) + 
  geom_line() + scale_color_manual(values = c(gmri_qual_extended[3],gmri_qual_extended[6],
                                              gmri_qual_extended[13],gmri_qual_extended[16])) + LGCtheme +
  labs(x="Banding Year", y="Total banded", 
       title="MALLs banded by month", color="Banding month")
```

Visualize # of bands by state south of 40N

```{r, fig.height=4.5, fig.width=10}
midcontinent_bands %>% 
  filter(gisb_lat<40) %>% 
  group_by(b_region_abbrev,b_month) %>% 
  summarise(n=n()) %>% 
  ggplot(aes(x=b_month, y=n,fill=as.factor(b_month))) + 
    facet_wrap(~forcats::fct_reorder(b_region_abbrev,n, .fun = sum, .desc = T), scales = "free_y", ncol = 7) +
  geom_bar(stat = "identity") + LGCtheme + theme(legend.position = "none") +
  scale_fill_manual(values = c(gmri_qual_extended[16],gmri_qual_extended[13],
                                              gmri_qual_extended[3],gmri_qual_extended[6])) +
  labs(x="Banding month", y="n banded", title="Pre-season banded MALLs south of Nebraska (<40N)")
```





```{r}
#Plot to confirm
ggplot(data = world) + geom_sf(fill="white") + 
  geom_sf(data=states, fill=NA) + 
  geom_sf(data=cent_flyway, fill="#6F7E4B") +
  geom_sf(data=miss_flyway, fill="#00608A") + 
  geom_point(data=midcontinent_bands, aes(x=gisb_long, y=gisb_lat)) +
  coord_sf(xlim = c(-130, -70), ylim = c(25, 70), expand = FALSE) + 
  labs(x="Banding longitude", y="Banding latitude")

```



