---
title: "Encounter processing"
author: "LGCarlson"
date: "7/30/2021"
output: html_document
---


## Encounter codes

_age:_

* 0 unknown

* 1 AHY

* 2 HY

* 3 Juvenile

* 4 Local

* 5 SY

* 6 ASY

_sex:_

* 0 unknown

* 4 M

* 5 F

_b_flyway / _r_flyway:_

* 2 Mississippi

* 3 Central

* 6 Canada

_b_region / r_region:_
[codes here](https://www.pwrc.usgs.gov/BBL/MANUAL/reg.cfm#region)
Canadian categories to keep for b_region: (US Flyways okay)

* 43 Northwest Territories

* 40 Nunavut

* 04 Alberta

* 79 Saskatchewan

* 45 Manitoba

* 68 Ontario

_b_month / r_month:_

* 1-12 or

* 82 Summer (06/21/YY thru 09/21/YY)

* 83 Spring (03/20/YY thru 06/20/YY)

* 92 Winter (12/21/YY-1 thru 03/19/YY)

* 93 Fall (09/22/YY thru 12/20/YY)

* 94 Hunting Season (09/01/YY thru 03/31/YY+1)

* 99 Unknown month and day of banding or encounter

_b_day / r_day:_

* 1-31 or

* 41 1st ten days in month

* 42 2nd ten days in month

* 43 3rd ten days in month

* 45 encountered during special hunting seasons

* 51-81 Encounter letter date or postmark date (plus 50)

* 99 Unknown day of banding or encounter

_how_obt:_

* 1 Shot

_hunt_season_surv:_

* number of hunting seasons survived (dbl)

_pres_cond:_

Categories to keep:

Code Bird condition  Band status

* 03	DEAD	UNKNOWN

* 04	DEAD	LEFT ON BIRD

* 05	DEAD	REMOVED

_status:_

* 2 transported

* 3 normal wild

* 4 hand reared

* 5 sick, injured, deformity

* 6,7,8 experimental, etc

_sp_num:_

* 1320 MALL

_ai_vai:_

* Categories to keep:

* Federal numbered metal band only. 

* Captured by spotlighting. 

* Misc. metal band (State, Provincial etc) with address or telephone number, plus Federal band.

* Miscellaneous band, Federal band, plus auxiliary marker(s). 


## Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(lubridate)
library(janitor)
library(rnaturalearth)
library(rnaturalearthdata)
library(ggspatial)
library(sf)
library(purrr)
```

```{r}
theme_set(theme_light())

#set personalized plotting theme
LGCtheme <- theme_light() + theme(panel.background = element_blank(), 
                panel.grid = element_blank(), 
                axis.text = element_text(size = 13), 
                axis.title = element_text(size = 14), 
                legend.text = element_text(size = 13), 
                legend.title = element_text(size = 14), 
                strip.background = element_blank(), 
                strip.text = element_text(size = 13, color = "black"))

gmri_qual_extended<-c("#003A54","#00608A","#5D96AF","#3F4B24","#6F7E4B","#97AA5C","#A8B212","#AF9700","#EACA00","#FFE961","#BB3F0E", "#EA4F12","#F4916B","#00736D", "#55A29E","#A0CDCB","#767676","#A7A7A7","#E2E2E2")
```



```{r}
#Create custom "not in" function
`%not_in%` <- purrr::negate(`%in%`)

# Midcontinent Canadian provinces
provs_keep<-c(43,40,04,79,45,68)

# Pacific or Atlantic Canadian provinces
provs_remove<-c(11,56,57,65,75,76,93)

# Band/capture types to keep
ai_vai_vec<-c("Federal numbered metal band only.","Captured by spotlighting.", "Misc. metal band (State, Provincial etc) with address or telephone number, plus Federal band.","Miscellaneous band, Federal band, plus auxiliary marker(s).")

# What to do with unknown encounter dates
unknown_date_tib<-tibble(r_day=c(1:31,41,42,43,51:81,99),r_day_adj=c(1:31,9,19,28,rep(28,32)))

# Name flyways
r_flyway_tib<-tibble(r_flyway=c(1,2,3,4,5,6,7), 
                     r_flyway_name=c("Atlantic","Mississippi","Central","Pacific","Alaska","Canada","Mexico"))

# Region (state/province) codes and names
r_region_codes<-read_csv(here::here("raw_data", "region_codes.csv")) %>% 
  filter(!is.na(region_abbrev)) %>% 
  dplyr::select(-region_name,-flyway_code) %>% 
  dplyr::rename("r_region" = "region_code","r_region_abbrev" = "region_abbrev")
```

## Load data

```{r, echo=FALSE, message=F, warning=F}
# Load in encounters from birds banded in Canada
ca_enc<-read_csv(here::here("raw_data", "encs_bandedin_canada.csv")) %>% 
  janitor::clean_names() %>% 
  mutate(b_flyway_name = "Canada") %>% 
  filter(b_region %in% provs_remove == F)
          #keep only data from birds banded in mid-continent Canadian provinces

# Load in encounters from birds banded in the Central flyway
cl_enc<-read_csv(here::here("raw_data", "encs_bandedin_central.csv")) %>% 
  janitor::clean_names() %>% 
  mutate(b_flyway_name = "Central") 

# Load in encounters from birds banded in the Miss flyway
ms_enc<-read_csv(here::here("raw_data", "encs_bandedin_miss.csv")) %>% 
  janitor::clean_names() %>% 
  mutate(b_flyway_name = "Mississippi") 
```


```{r}
# Initial data check

#should all be 6
table(ca_enc$b_flyway)

#should all be 3
table(cl_enc$b_flyway)

#should all be 2
table(ms_enc$b_flyway)

#can be any flyway
table(ca_enc$r_flyway);table(cl_enc$r_flyway);table(ms_enc$r_flyway)
```


```{r}
#Join all encounters 

all_encs<-bind_rows(ca_enc, cl_enc, ms_enc) %>% 
  left_join(r_flyway_tib, by = "r_flyway")
```



## Check data

```{r}
#should all be 1
table(all_encs$how_obt) 

# will need to remove years > 2011
table(all_encs$r_year)

# will need to remove (if unknown) or specify (approx known) non-real months
table(all_encs$r_month)

# will need to remove (if unknown) or specify (approx known) non-real days
table(all_encs$r_day)

# will need to remove (if unknown) or specify (approx known) non-real days
table(all_encs$b_day)
```


## Processing code

_Modifications (in order of operation):_

* remove indv. where banding month is unknown

* remove indv. where recovery month is unknown

* if b_day is unknown (99), assign b_day = 28

* join unknown_date_tib by r_day, so that r_day_adj can be used when recovery month is known, but recovery day is unknown

* create a new column with banding date (b_date)

* create a new column with recovery date (r_date)

* create a new column with recovery date in julian days (r_doy)

* limit years to 1950-2020

* keep only pre-season banded birds (banding months 6-9)

* keep only normal, wild birds  (status = 3)

* keep only known sexes (sex = 4,5)

* keep only desired ai_vai's

* remove birds for which hss cannot be calculated (hunt_season_survived = 99)

* re-order columns manually

* remove 3 weird duplicate bands

```{r}
final_encs<-all_encs %>% 
  filter(b_month <= 12, 
         r_month <= 12) %>% 
  mutate(b_day = ifelse(b_day == 99, 28, b_day)) %>% 
  left_join(unknown_date_tib, by="r_day") %>% 
  mutate(b_date = as.Date(paste(b_year,b_month,b_day, sep = "-"))) %>% 
  mutate(r_date = as.Date(paste(r_year,r_month,r_day_adj, sep = "-"))) %>% 
  mutate(r_doy = lubridate::yday(r_date)) %>% 
  filter(b_year >=1950 & b_year <=2020) %>% 
  filter(b_month >=6 & b_month <=9) %>% 
  filter(r_month %in% c(9,10,11,12,1,2)) %>% 
  filter(ifelse(r_month == 2, r_doy < 75, r_day < 366)) %>% 
  filter(status == 3) %>% 
  filter(sex == 4 | sex == 5) %>% 
  filter(ai_vai %in% ai_vai_vec) %>% 
  filter(hunt_season_surv != 99) %>% 
  left_join(r_region_codes, by = c("b_region" = "r_region")) %>% 
  dplyr::rename("b_region_abbrev" = "r_region_abbrev") %>% 
  left_join(r_region_codes, by = "r_region") %>% 
  dplyr::select(band,age,sex,sp_num,how_obt,pres_cond,status,hunt_season_surv,b_flyway,b_flyway_name,
                b_date,b_year,b_month,b_day,b_region,b_region_abbrev,b_dir,b_coord_precision,
                b_lat,b_long,gisb_lat,gisb_long,
                r_flyway,r_flyway_name,r_date,r_year,r_month,r_day,r_region,r_region_abbrev,r_dir,
                r_coordinate_precision,r_lat,r_long,gisr_lat,gisr_long,ai_vai,everything())  %>% 
  filter(!is.na(gisr_long)) %>% 
  filter(!is.na(gisr_lat)) %>% 
  filter(band %not_in% c("135748061","149776745","236737635"))

```




```{r}
#confirm that "band" field works as a unique identifier (it should) 
length(final_encs$band)

length(unique(final_encs$band))
```


## Spatial processing

```{r, message=F, warning=F, echo=F}
# load in world and state sf
world <- rnaturalearth::ne_countries(scale = "medium", returnclass = "sf")

states <- rnaturalearth::ne_states(country = c("canada", "united states of america"), returnclass = "sf")


#load in flyway shapefiles
cent_flyway <- sf::st_read(here::here("shapefiles","Central_Flyway.shp")) %>% 
  st_transform(crs = "+init=epsg:4326") 

miss_flyway <- sf::st_read(here::here("shapefiles","Mississippi_Flyway.shp")) %>% 
  st_transform(crs = 4326) 

pac_flyway <- sf::st_read(here::here("shapefiles","Pacific_Flyway.shp")) %>% 
  st_transform(crs = 4326) 

atl_flyway <- sf::st_read(here::here("shapefiles","Atlantic_Flyway.shp")) %>% 
  st_transform(crs = 4326) 
```

Plot banding locations. Will need to remove birds banded in the Pacific flyway (outside mid-continent). 

```{r}
ggplot(data = world) + geom_sf(fill="white") + 
  geom_sf(data=states, fill=NA) + 
  geom_sf(data=cent_flyway, fill="#6F7E4B") +
  geom_sf(data=miss_flyway, fill="#00608A") + 
  geom_point(data=final_encs, aes(x=gisb_long, y=gisb_lat),size=0.85) +
  coord_sf(xlim = c(-130, -70), ylim = c(25, 70), expand = FALSE) +
  labs(x="Banding longitude", y="Banding latitude", title="Banding locations (full data)")
```

Remove encounters of birds that were *banded* in Pacific flyway. Can be *encountered* anywhere still.

```{r}
#Create spatial data
final_encs_sf<-st_as_sf(final_encs, coords = c("gisb_long","gisb_lat"), crs = 4326) 
class(final_encs_sf)

length(unique(final_encs_sf$band))
```

```{r}
#This selects all individuals that were banded inside the Pacific flyway shapefile
banded_inpac <- final_encs_sf[pac_flyway, ]

#Creates a vector of Pacific flyway banded individuals
enc_to_remove<-c(banded_inpac$band)

#Creates dataframe of Pacific flyway banded individuals to plot and check for correctness
final_bands_inpac<-final_encs %>% 
  filter(band %in% enc_to_remove)

#Plot to confirm
ggplot(data = world) + geom_sf(fill="white") + 
  geom_sf(data=states, fill=NA) + 
  geom_sf(data=pac_flyway, fill="#F4916B") +
  geom_point(data=final_bands_inpac, aes(x=gisb_long, y=gisb_lat),size=0.85) +
  coord_sf(xlim = c(-130, -70), ylim = c(25, 70), expand = FALSE) + 
  labs(x="Banding longitude", y="Banding latitude", title="Banding locations (to remove)")
```

```{r}
#Remove individuals banded in Pacific flyway
mcb_encs<-final_encs %>% 
  filter(band %not_in% enc_to_remove)

#Plot to confirm
ggplot(data = world) + geom_sf(fill="white") + 
  geom_sf(data=states, fill=NA) + 
  geom_sf(data=cent_flyway, fill="#6F7E4B") +
  geom_sf(data=miss_flyway, fill="#00608A") + 
  geom_point(data=mcb_encs, aes(x=gisb_long, y=gisb_lat),size=0.85) +
  coord_sf(xlim = c(-130, -70), ylim = c(25, 70), expand = FALSE) + 
  labs(x="Banding longitude", y="Banding latitude", title="Banding locations (filtered data)")

```

```{r}
# Length of filtered dataset
length(mcb_encs$band)

# Banded in midcontinent + banded in Pacific should = length of original dataframe
length(mcb_encs$band) + length(final_bands_inpac$band)

length(final_encs$band)
#It does!
```


# Spatial categorizing

```{r}
# Create new column to categorize regions midcontinent US, midcontinent Canada, and other

mcb_encs$r_region_filter = NA

mcb_encs$r_region_filter[mcb_encs$r_flyway_name == "Central"] = "Midcontinent US"

mcb_encs$r_region_filter[mcb_encs$r_flyway_name == "Mississippi"] = "Midcontinent US"

mcb_encs$r_region_filter[mcb_encs$r_flyway_name == "Pacific"] = "Other"

mcb_encs$r_region_filter[mcb_encs$r_flyway_name == "Atlantic"] = "Other"

mcb_encs$r_region_filter[mcb_encs$r_region_abbrev == "AK"] = "Other"

mcb_encs$r_region_filter[mcb_encs$r_region_abbrev == "MX"] = "Other"

mcb_encs$r_region_filter[mcb_encs$r_flyway_name == "Canada"] = "Other"

mcb_encs$r_region_filter[mcb_encs$r_region_abbrev == "ON"] = "Midcontinent Canada"

mcb_encs$r_region_filter[mcb_encs$r_region_abbrev == "MB"] = "Midcontinent Canada"

mcb_encs$r_region_filter[mcb_encs$r_region_abbrev == "SK"] = "Midcontinent Canada"

mcb_encs$r_region_filter[mcb_encs$r_region_abbrev == "AB"] = "Midcontinent Canada"


table(mcb_encs$r_region_filter)

# There are 8 encounters that don't have a recovery region associated with them (even though they do have real GPS points) so will not be categorized correctly (...consider deleting?)

mcb_encs %>% 
  filter(is.na(r_region_filter)==T)
```


```{r, fig.height=4, fig.width=8}
#Visualize encounter locations and check region assignments
ggplot(data = world) + geom_sf(fill=NA) + 
  geom_sf(data=states, fill=NA) + 
  geom_point(data=mcb_encs, aes(x=gisr_long, y=gisr_lat, color=r_region_filter),size=0.85) +
  coord_sf(xlim = c(-130, -70), ylim = c(25, 70), expand = FALSE) + 
  labs(x="Encounter longitude", y="Encounter latitude", color="Recovery region", title="Encounter location")
```


## Save new datafile

```{r}
## write_csv(mcb_encs, here::here("proc_data","encounter_data.csv"))
```


## Spatial visualization 


```{r, fig.height=8, fig.width=6}
#Visualize encounter locations
ggplot(data = world) + geom_sf(fill=NA) + 
  geom_sf(data=states, fill=NA) + 
  geom_sf(data=cent_flyway, fill="#6F7E4B",alpha=0.6) +
  geom_sf(data=miss_flyway, fill="#00608A",alpha=0.6) + 
  geom_point(data=mcb_encs, aes(x=gisr_long, y=gisr_lat),size=0.85) +
  coord_sf(xlim = c(-130, -70), ylim = c(25, 70), expand = FALSE) + 
  labs(x="Encounter longitude", y="Encounter latitude")
```



## Visualize data

Majority of encountered individuals encountered in the midcontinent were banded in Canada.
More males encountered than females. 

```{r, fig.height=4, fig.width=10}
mcb_encs %>% 
  group_by(r_year, sex, b_flyway_name) %>% 
  summarise(total_encountered = n()) %>% 
  ggplot(aes(x=r_year, y=total_encountered, color=as.factor(sex))) + geom_line() + 
  facet_wrap(~b_flyway_name, ncol=3) + 
  scale_color_manual(values=c("#5D96AF","#C07F9C")) + LGCtheme + theme(legend.position = "none") + 
  labs(x="Encounter Year", y="Total encountered", 
       title="MALLs encountered by banding flyway, by sex")
```

Majority of individuals are encountered in Midcontinent US

```{r, fig.height=4, fig.width=10}
mcb_encs %>% 
  filter(r_flyway > 0 & is.na(r_flyway) == F) %>% 
  group_by(r_year, b_flyway_name, r_region_filter) %>% 
  summarise(total_encountered = n()) %>% 
  ggplot(aes(x=r_year, y=total_encountered, color=b_flyway_name)) + geom_line() + 
  facet_wrap(~forcats::fct_reorder(r_region_filter,total_encountered, .fun = sum, .desc = T), ncol=3) + 
  scale_color_manual(values=c(gmri_qual_extended[13],gmri_qual_extended[5],gmri_qual_extended[2])) + 
  LGCtheme + theme(legend.position = "bottom") + 
  labs(x="Encounter Year", y="Total encountered", color="Banding flyway",
       title="MALLs encountered in each region, by banding flyway")
```

Most individuals were ecountered during Oct, Nov, Dec

```{r, fig.height=4.5, fig.width=7.5}
mcb_encs %>% 
  mutate(r_month = as.factor(r_month)) %>% 
  group_by(r_year, r_month) %>% 
  summarise(total_encountered = n()) %>% 
  ggplot(aes(x=r_year, y=total_encountered, color=forcats::fct_reorder(r_month, total_encountered, .desc = T))) + 
  geom_line() + scale_color_manual(values = gmri_qual_extended) + LGCtheme +
  labs(x="Encounter Year", y="Total encountered", 
       title="MALLs encountered by month", color="Encounter month")

```


More mid-continent banded mallards and encountered in Arkansas than in any other state [wenowdis](https://www.youtube.com/watch?v=2kQxVwYwrME)

```{r, fig.height=8, fig.width=5.25}
mcb_encs %>% 
  group_by(r_region_abbrev, r_flyway_name) %>% 
  summarise(total_encountered = n()) %>% 
  filter(!is.na(r_region_abbrev), 
         r_region_abbrev != "Unknown") %>% 
  ggplot(aes(x=forcats::fct_reorder(r_region_abbrev,total_encountered,.fun = sum, .desc = F), 
          y=total_encountered, fill=forcats::fct_reorder(r_flyway_name,total_encountered,.fun = sum, .desc = T))) + 
  geom_bar(stat = "identity") + coord_flip() +
  scale_fill_manual(values=c(gmri_qual_extended[2],gmri_qual_extended[5],gmri_qual_extended[13],
                    gmri_qual_extended[16],gmri_qual_extended[10],gmri_qual_extended[7],gmri_qual_extended[11])) + 
  LGCtheme + theme(axis.text.y = element_text(size=9, angle=0)) +
  labs(x="Postal abbreviation",y="Total encountered", title="MALLs encountered by state",fill="Region flyway") +
  theme(legend.position = c(0.8,0.18))

```


```{r, fig.height=8, fig.width=6}
#Visualize encounter locations
ggplot(data = world) + geom_sf(fill=NA) + 
  geom_sf(data=states, fill=NA) + 
  geom_sf(data=cent_flyway, fill="lightgrey",alpha=0.6) +
  geom_sf(data=miss_flyway, fill="lightgrey",color=NA,alpha=0.6) + 
  geom_point(data=mcb_encs, aes(x=gisr_long, y=gisr_lat),size=0.85) +
  coord_sf(xlim = c(-130, -65), ylim = c(23, 68), expand = FALSE) + 
  labs(x="Encounter longitude", y="Encounter latitude")
```


## Additional mapping

```{r}
enc_flyways_sf<-st_as_sf(mcb_encs, coords = c("gisr_long","gisr_lat"), crs = 4326) 
class(enc_flyways_sf)

recv_inpac <- enc_flyways_sf[pac_flyway, ]
recv_inatl <- enc_flyways_sf[atl_flyway, ]  
recv_incent <- enc_flyways_sf[cent_flyway, ]
recv_inmiss <- enc_flyways_sf[miss_flyway, ]

pac<-tibble(band=c(recv_inpac$band)) %>% 
  mutate(r_flyway_fws = "Pacific")
atl<-tibble(band=c(recv_inatl$band)) %>% 
  mutate(r_flyway_fws = "Atlantic")
cent<-tibble(band=c(recv_incent$band)) %>% 
  mutate(r_flyway_fws = "Midcontinent")
miss<-tibble(band=c(recv_inmiss$band)) %>% 
  mutate(r_flyway_fws = "Midcontinent")

add_r_flyways<-bind_rows(pac,atl,cent,miss) 

mcb_box<-mcb_encs %>% 
  left_join(add_r_flyways, by = "band") %>% 
  mutate(gisb_latbox = 1 * gisb_lat %/% 1) %>% 
  mutate(gisb_longbox = 2 * gisb_long %/% 2) 

mcb_encs %>% 
  left_join(add_r_flyways, by = "band") %>% 
  filter(is.na(r_flyway_fws))
```










```{r}
enc1x2<-mcb_encs %>% 
  mutate(gisr_latbox = 1 * gisr_lat %/% 1) %>% 
  mutate(gisr_longbox = 2 * gisr_long %/% 2) %>% 
  group_by(gisr_latbox, gisr_longbox) %>% 
  summarise(n=n())
```


```{r}
ggplot(data=enc1x2) + geom_sf(data = world,fill=NA) + 
  geom_sf(data=states, fill=NA) + 
  geom_sf(data=cent_flyway, fill="lightgrey",alpha=0.6) +
  geom_sf(data=miss_flyway, fill="lightgrey",color=NA,alpha=0.6) + 
  geom_point(data=enc1x2, aes(x=gisr_longbox, y=gisr_latbox,size=n)) + 
  coord_sf(xlim = c(-130, -65), ylim = c(23, 68), expand = FALSE) + 
  labs(x="Encounter longitude", y="Encounter latitude")
```

```{r}

mex<-c("Jalisco","Aguascalientes","Nayarit","Guanajuato","Queretaro de Arteaga","Hidalgo","Michoacan de Ocampo","Mexico","Distrito Federal","Colima","Morelos","Yucatan","Campeche","Puebla","Quintana Roo","Tlaxcala","Guerrero","Oaxaca","Chiapas", "Tabasco","Chihuahua","Coahuila De Zaragoza","Durango","Zacatecas","San Luis Potosi","Nuevo Leon","Tamaulipas","Veracruz-Llave")

cent_nomex<-cent_flyway %>% filter(Area_Name %not_in% mex)
```


```{r}

cent_flyway <- sf::st_read("D:/UAM/GIS_Layers/Central_Flyway.shp") %>% 
  st_transform(crs = "+init=epsg:4326")  %>% 
  summarise()

miss_flyway <- sf::st_read("D:/UAM/GIS_Layers/Mississippi_Flyway.shp") %>% 
  st_transform(crs = 4326) %>% 
  summarise()


lakes <- st_read("D:/UAM/GIS_Layers/ne_10m_lakes.shp") %>% 
  filter(name %in% c("Lake Erie","Lake Huron", "Lake Michigan", "Lake Ontario", "Lake Superior"))
```


```{r}
enc1x1<-mcb_encs %>% 
  mutate(gisr_latbox = 1 * gisr_lat %/% 1) %>% 
  mutate(gisr_longbox = 1 * gisr_long %/% 1) %>% 
  group_by(gisr_latbox, gisr_longbox) %>% 
  summarise(n=n()) 

hist(enc1x1$n)
```


```{r, fig.height=6,fig.width=7}
ggplot(data=enc1x1) + 
  geom_tile(data=enc1x1, aes(x=gisr_longbox, y=gisr_latbox,fill=n)) + 
  scale_fill_gradient(low="#ededed", high = "black",breaks = c(0, 2000, 6000, 10000,14000,18000)) +
  geom_sf(data=cent_flyway, fill=NA,alpha=0.6,size=0.3) +
  geom_sf(data=miss_flyway, fill=NA,alpha=0.6,size=0.3) + 
  geom_sf(data = world,fill=NA,size=0.3) + 
  coord_sf(xlim = c(-130, -65), ylim = c(23, 67), expand = FALSE) + 
  scale_x_continuous(breaks=c(-120,-100,-80)) + 
  scale_y_continuous(breaks=c(30,45,60)) + 
  theme(panel.grid = element_blank()) + 
  guides(fill = guide_colourbar(ticks = T, barwidth = 0.75, barheight = 6, nbin = 6)) +
  labs(x="", y="",fill="Encounters (n)") +
  theme(axis.text = element_text(size=11), axis.text.y = element_text(angle=90,hjust = 0.5))


#ggsave("recoveries_heatmap.pdf", plot=last_plot(), height = 5, width = 6)
```




```{r, fig.height=6,fig.width=7}
ggplot(data=mcb_encs) + 
  geom_sf(data = world,fill=NA,size=0.3) + 
  geom_sf(data=cent_flyway, fill=NA,alpha=0.6,size=0.3) +
  geom_sf(data=miss_flyway, fill=NA,alpha=0.6,size=0.3) + 
  geom_point(data=mcb_encs, aes(x=gisr_long, y=gisr_lat),size=0.75, alpha=0.25) +
  theme(panel.grid = element_blank()) + 
  labs(x="", y="") +  
  coord_sf(xlim = c(-130, -65), ylim = c(23, 67), expand = FALSE) + 
  scale_x_continuous(breaks=c(-120,-100,-80)) + 
  scale_y_continuous(breaks=c(30,45,60)) + 
  theme(axis.text = element_text(size=11), axis.text.y = element_text(angle=90,hjust = 0.5))


#ggsave("recoveries_points.pdf", plot=last_plot(), height = 5, width = 6)
```



```{r}
enc1x1$labels = NA
enc1x1$labels[enc1x1$n <=500] = "1-500"
enc1x1$labels[enc1x1$n > 500] = "501-1000"
enc1x1$labels[enc1x1$n >1000] = "1001-5000"
enc1x1$labels[enc1x1$n >5000] = "5000-10000"
enc1x1$labels[enc1x1$n >10000] = ">10000"

enc1x1$breaks = NA
enc1x1$breaks[enc1x1$n <=500] = 1
enc1x1$breaks[enc1x1$n > 500] = 2
enc1x1$breaks[enc1x1$n >1000] = 3
enc1x1$breaks[enc1x1$n >5000] = 4
enc1x1$breaks[enc1x1$n >10000] = 5

enc1x1$breaks = as.character(enc1x1$breaks)
```


```{r}
ggplot(data=mcb_encs) + 
  geom_sf(data = world,fill=NA,size=0.2) + 
  geom_sf(data=cent_flyway, fill=NA,alpha=0.6,size=0.2) +
  geom_sf(data=miss_flyway, fill=NA,alpha=0.6,size=0.2) + 
  geom_point(data=enc1x1, aes(x=gisr_longbox, y=gisr_latbox,size=n,color=breaks)) +
  scale_size_continuous(range=c(0.25,7), breaks = c(500,1000,5000,10000,15000), 
                        labels = c("1 - 500","501 - 1,000","1,001 - 5,000","5,001 - 10,000","> 10,000")) +
  scale_color_manual(values=rev(c("gray20", "gray30", "gray50","gray70", "gray90"))) +
  theme(panel.grid = element_blank()) + 
  #guides(color = FALSE) +
  labs(x="", y="",size="Recoveries") +  
  coord_sf(xlim = c(-130, -64), ylim = c(20, 65), expand = FALSE) + 
  scale_x_continuous(breaks=c(-120,-100,-80)) + 
  scale_y_continuous(breaks=c(30,45,60)) + 
  theme(axis.text = element_text(size=11), axis.text.y = element_text(angle=90,hjust = 0.5))


#ggsave("recoveries_points_sizecolor2.pdf", plot=last_plot(), height = 8, width = 8)
```


```{r}
ggplot(data=mcb_encs) + 
  geom_sf(data = world,fill=NA,size=0.3) + 
  geom_sf(data=cent_flyway, fill=NA,alpha=0.6,size=0.3) +
  geom_sf(data=miss_flyway, fill=NA,alpha=0.6,size=0.3) + 
  geom_point(data=enc1x1, aes(x=gisr_longbox, y=gisr_latbox,size=n,color=breaks)) +
  scale_size_continuous(range=c(0.25,7)) +
  theme(panel.grid = element_blank()) + 
  labs(x="", y="",size="Encounters (n)") +  
  coord_sf(xlim = c(-130, -65), ylim = c(23, 67), expand = FALSE) + 
  scale_x_continuous(breaks=c(-120,-100,-80)) + 
  scale_y_continuous(breaks=c(30,45,60)) + 
  theme(axis.text = element_text(size=11), axis.text.y = element_text(angle=90,hjust = 0.5))


#ggsave("recoveries_pointsize.pdf", plot=last_plot(), height = 5, width = 6)
```


